### 2022-10-25 README.Rmd
some bug fixes
also added more plots for the markdown file to reproduce results from paper

This Repository contains the Bremla package for uncertainty quantification and synchronization for layer-counted proxy archives, including the data and code to reproduce the results for

@@

Warning: The provided real data example generates several thousand simulated chronologies with individual lengths exceeding 18,000. More if synchronized chronologies is also computed. This will require a significant amount of memory. If needed, reduce the number of samples generated or free up memory before use.

## Example
This is a real data example which produces 5000 simulations from a synchronized time scale.
```{r example, message=FALSE, warning=FALSE, include=TRUE}
require(INLA)
library(bremla)
data("event_intervals")
data("events_rasmussen")
data("NGRIP_5cm")

>>>>

This repository contains the bremla package (**B**ayesian **Re**gression **M**odeling of **L**ayer-counted **a**rchives) for uncertainty quantification and synchronization for layer-counted proxy archives, including the data and code to reproduce the results for

@@

Warning: The provided real data example generates several thousand simulated chronologies with individual lengths exceeding 18,000. More if both synchronized and unsynchronized chronologies computed. This will require a significant amount of memory. If needed, reduce the number of samples generated or free up memory before use.

## Example
This is a real data example for the GICC05 chronology of the NGRIP ice core record which produces 5000 simulations from a time scale synchronized using the Adolphi tie-points.
```{r example, message=FALSE, warning=FALSE, include=TRUE}
require(INLA)
library(bremla)
data("event_intervals")
data("events_rasmussen")
data("NGRIP_5cm")

### 2022-10-25 inst/Examples/Paper_results/Paper_results_2.Rmd
some bug fixes
also added more plots for the markdown file to reproduce results from paper

>>>>

## Illustration example

The data used to generate the illustrative examples in the paper is as follows.

```{r illustrationexample}
rm(list=ls())
require(stats)
library(ggplot2)
set.seed(123)
n=800
y0 = 0
z0 = 0
phi=0.8; sigma=1.2
depth = z0+seq(from=1,to=n,length.out=n)

covariate = arima.sim(model=list(ar=c(0.99)),n=n)
noise = sigma*arima.sim(model=list(ar=c(phi)),n=n,sd=sqrt(1-phi^2))

dy = 10 + depth*5/n+covariate+noise
y = y0+cumsum(dy)

data=data.frame(age=c(y0,y),depth=c(z0,depth),dy=c(0,as.numeric(dy)),covariate=c(0,covariate))
#data=rbind(c(y0,z0,0,0),df)
formula=dy~ 1 + depth + covariate 
nsims = 10000

# Unsynchronized
res = bremla(formula,data,reference.label="simulated time scale",nsims=nsims,
             control.fit=list(noise="ar1"),
             x.label="Depth",
             y.label="Ensemble age - observed age",
             control.sim = list(synchronized=FALSE,
                                nsims=nsims))

library(ggplot2)
orichron = res$data$age
ggd = data.frame(depth=res$data$depth,
                 mean = res$simulation$summary$mean-orichron,
                 lower = res$simulation$summary$lower-orichron,
                 upper = res$simulation$summary$upper-orichron)
ggp = ggplot(data=ggd,aes(depth))+theme_bw()+xlab("Depth")+ylab("Ensemble age - observed age")+
  geom_ribbon(aes(ymin=lower,ymax=upper),col="red",fill="red",alpha=0.3,size=0.7)+
  geom_line(aes(y=mean),col="blue",size=0.7)
# plot(res)
#ggsave("illustrative-unsync.eps",device=cairo_ps,width=4800,height=2000,units="px")

# Fixed tie-points
ntie = 3
tiesims = matrix(NA,nrow=nsims,ncol=ntie)
tiedepths = c(100,300,600)
tieoffset = c(-50,0,100)
#tieage = c(450,3000,8500)
tieage = y[tiedepths]+tieoffset

for(i in 1:ntie){
  tiesims[,i] = rep(tieage[i],nsims)
}

res2 = tiepointsimmer(res,synchronization = list(locations=tiedepths,locations_unit="depth",
                       samples = tiesims)
               )

res2 = bremla_synchronized_simulation(res2,control.sim=list(synchronized=TRUE))

ggd = data.frame(depth=res2$data$depth,
                 mean = res2$simulation$summary_sync$mean-orichron,
                 lower = res2$simulation$summary_sync$lower-orichron,
                 upper = res2$simulation$summary_sync$upper-orichron)

oritie = orichron[tiedepths]
ggd2 = data.frame(depth=res2$tie_points$locations,
                  mean=res2$simulation$summary_sync$mean[tiedepths]-oritie)

ggp = ggplot(data=ggd,aes(depth))+theme_bw()+xlab("Depth")+ylab("Ensemble age - observed age")+
  geom_ribbon(aes(ymin=lower,ymax=upper),col="red",fill="red",alpha=0.3,size=0.7)+
  geom_line(aes(y=mean),col="blue",size=0.7)+
  geom_point(data=ggd2,mapping=aes(x=depth,y=mean),col="magenta",size=1.2)
#plot(res)
#ggsave("illustrative-fixedsync.eps",device=cairo_ps,width=4800,height=2000,units="px")

tieoffset = c(-50,0,100)
tiemeans = y[tiedepths]+tieoffset
#tiemeans = c(450,3000,8500)
tiesds = c(10,20,30)
res3 = tiepointsimmer(res,synchronization = list(locations=tiedepths,
                                                locations_unit="depth",
                                                method="normal",
                                                params=list(mean=tiemeans,
                                                            sd=tiesds))
               )
res3 = bremla_synchronized_simulation(res3,control.sim=list(synchronized=TRUE))
ggd = data.frame(depth=res3$data$depth,
                 mean = res3$simulation$summary_sync$mean-orichron,
                 lower = res3$simulation$summary_sync$lower-orichron,
                 upper = res3$simulation$summary_sync$upper-orichron)

oritie = orichron[tiedepths]
ggd2 = data.frame(depth=res3$tie_points$locations,
                  lower=res3$simulation$summary_sync$lower[tiedepths]-oritie,
                  upper=res3$simulation$summary_sync$upper[tiedepths]-oritie)

ggp = ggplot(data=ggd,aes(depth))+theme_bw()+xlab("Depth")+ylab("Ensemble age - observed age")+
  geom_ribbon(aes(ymin=lower,ymax=upper),col="red",fill="red",alpha=0.3,size=0.7)+
  geom_line(aes(y=mean),col="blue",size=0.7)+
  geom_segment(data=ggd2,mapping=aes(x=depth,xend=depth,y=lower,yend=upper), 
               col="magenta",size=0.7)

#plot(res3)
#ggsave("illustrative-sync.eps",device=cairo_ps,width=4800,height=2000,units="px")

```
